---
title: "Tables and Figures DLNM"
output: 
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# Load necessary libraries
library(readxl)     # reading Excel files
library(tidyverse)      # data manipulation
library(labelled)   # labeling columns 
library(here) # file paths
library(haven)
library(stargazer)
library(dlnm)
library(data.table)
library(forcats) # For reordering factor levels
library(broom)
library(MASS)
library(brant) # test proportional odds assumption for ordinal logistic regression
library(gt)
library(patchwork)
# source in code to load and clean data

## override mass::select and use dplyr as default
select <- dplyr::select 

source(here("data_cleaning", "clean_stress_data.R")) # pss_recode and crisis_recode are the output datasets of interest



# load selected temp data
era5 <- read_csv(here("data", "GRAPHS-ERA5-CHIRPS", "GRAPHS-ERA5-Heatstress-Tw-1991missing.csv")) %>%
  filter(date > as.Date("2013-01-01")) %>%
  mutate(community = case_when(community == "AKORA/AT" ~ "AKORA",
                               TRUE ~ community)) %>%
  mutate(tw_delta = max_tw - min_tw)

# load in precip data 
precip <- read_csv(here("data", "GRAPHS-ERA5-CHIRPS", "GRAPHS-CHIRPS-Precip.csv")) 

precip_clean <- precip %>%
  select(community, date, precip) %>%
  filter(date > as.Date("2013-01-01")) %>%
  mutate(date = lubridate::ymd(date),
         month = lubridate::month(date))%>%
  mutate(community = case_when(community == "AKORA/AT" ~ "AKORA",
                               TRUE ~ community))

# check definition of rainy season 
precip_summary <- precip_clean %>%
  group_by(month) %>%
  summarise(total_precip = sum(precip))


# load selected temp data
tw_daily_max <- era5 %>%
  dplyr::select(community, date, max_tw)  # select relevant columns

tw_daily_mean <- era5 %>%
  dplyr::select(community, date, mean_tw)  # select relevant columns

tw_daily_min <- era5 %>%
  dplyr::select(community, date, min_tw)  # select relevant columns

```



## Sum NDS Category


```{r, echo=FALSE, message = FALSE, warning=FALSE}
### LOAD THE CRYSIS DATA

## CRYSIS data
crisis_simple <- crisis_recode %>%
  mutate(
    quessetdt = sprintf("%04d", quessetdt),  # Ensure quassetdt has 4 digits
    quessetdt = sub("(\\d{2})(\\d{2})", "\\1:\\2", quessetdt),  # Insert colon
    survey_datetime = as.POSIXct(paste(quessetd, quessetdt), format = "%Y-%m-%d %H:%M")) %>%
    select(mstudyid, fin_events:survey_datetime, neg_fin_groups, neg_rel_groups, neg_home_groups, datdeliv:survey_pre_post_birth, gestage_days, date_conception, married, wealthindex, age, bmi, medlev, fan, pregchn) %>%
  select(-c(crireadneg:crikidneg)) %>%
  mutate(month_delivery = lubridate::month(datdeliv),
         season_delivery = case_when(month_delivery %in% c(4,5,6,7,8,9,10) ~ "wet", 
                                     month_delivery %in% c(1,2,3,11,12) ~"dry",
                                     TRUE ~ NA)) %>%
  
  mutate(married_bin = case_when(
      married %in% c(1, 2) ~ "partnered",
      married %in% c(3, 4, 5, 6) ~ "not_partnered",
      TRUE ~ NA_character_
    ),
    married_bin = factor(married_bin)) %>%
  
  mutate(medlev_bin = case_when(
      medlev == 1 ~ "None",
      medlev == 2 ~ "Primary",
      medlev %in% c(3, 4) ~ "≥ JHS",
      TRUE ~ NA_character_
    ),
    medlev_bin = factor(medlev_bin)) %>%
  
  mutate(survey_date = as.Date(survey_datetime))  # Extract the date part of survey_datetime


```


### Max tw

```{r, echo=FALSE, message = FALSE, warning=FALSE}

## Create Daily Lag Frame:
max_gestation_days <- 210 ## Two weeks into third trimester. This was chosen to avoid post-survey exposure while keeping the full second trimester and sample size. Went two weeks into third trimester so estimates in second trimester more stable, as DLNMs uncertain at edge periods. 

pre_conception_days <- 30

crisis_dlnm_eligible <- crisis_simple %>% 
  filter(gestage_at_survey >= max_gestation_days) %>% # filter 
  filter(!is.na(bmi))


# Function to process temperature data for the selected metric
process_temperature_data <- function(era5_data, metric) {
  era5_data %>%
    select(community, date, !!sym(metric)) %>%
    rename(temp_metric = !!sym(metric)) %>%
    distinct()
}

# Process temp data
temp_data_clean <- process_temperature_data(era5_data = era5, metric = "max_tw")



# Generate lags relative to conception
create_conception_aligned_lags <- function(df, temp_df) {
  df %>%
    select(mstudyid, vname, date_conception) %>%
    rowwise() %>%
    mutate(day_relative = list(-pre_conception_days:max_gestation_days),
           date = list(date_conception + day_relative)) %>%
    unnest(c(day_relative, date)) %>%
    left_join(temp_df, by = c("vname" = "community", "date" = "date"))  # join temp
}


temp_lags_long <- create_conception_aligned_lags(crisis_dlnm_eligible, temp_data_clean)

# Step 4: Pivot to wide format (like DLNM expects)
temperature_lags_wide <- temp_lags_long %>%
  select(mstudyid, date_conception, day_relative, temp_metric) %>%
  pivot_wider(
    names_from = day_relative,
    values_from = temp_metric,
    names_prefix = "day_"
  ) %>%
  left_join(crisis_dlnm_eligible %>% select(mstudyid, sum_nds_category, neg_fin_groups, neg_rel_groups, neg_home_groups, age, bmi, wealthindex, pregchn, season_delivery, married_bin, medlev_bin), by = "mstudyid") %>%
  filter(!is.na(day_0))

temp_data_matrix <- temperature_lags_wide %>%
  select(starts_with("day_")) %>%
  as.matrix()

```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# set up DLNM model parameters, fit model, and get predictions

## SET UP PARAMS

# predictor DLNM model parameters -- captures the relationship between the predictor variable (temp) and the outcome
var_arg = list(fun = "ns", df = 4) # smooth cubic polynomial (degree = 3) between the boundary knots

# lag DLNM model parameters -- captures the lagged effects of the predictor variable (temp) over time.
lag_arg = list(fun = "ns", df = 3) # smooth cubic polynomial (degree = 3) between the boundary knots 

# cross-basis matrix for DLNM
cb = crossbasis(temp_data_matrix, lag = c(0, ncol(temp_data_matrix) - 1), argvar = var_arg, arglag = lag_arg) # basis for the predictor and lag 

# ordered logistic regression model
model_polr <- polr(
  formula = sum_nds_category ~ cb + age + bmi + wealthindex + season_delivery + pregchn + married_bin + medlev_bin,
  data = temperature_lags_wide,
  method = "logistic" 
)


# GET PREDICTIONS (in reference to median temp)
temp_median <- round(median(as.vector(temp_data_matrix)), 1) # centering value (median). 
crysis_temp_median <- temp_median

# generate predictions for a DLNM, set centering value as the median of all temperatures
cpred = crosspred(cb, 
                  model_polr, 
                  ci.level = 0.95, # 95% confidence intervals calculated around predicted values
                  cen = temp_median, # predictions are relative to this baseline temperature.
                  by = 0.1) # predictions are generated at intervals of 0.1 units of temperature



png(here("outputs", "contour_max_tw.png"), width = 8, height = 6, units = "in", res = 300)

par(mar = c(4.8, 5.6, 2.2, 6.2))  # bottom, left, top, right

plot(cpred, "contour",
     xlab = "Max temperature (°C)",
     ylab = "Days relative to conception (0 = 30 days preconception)",
     key.title = title(main = "Change in \nlog odds", cex.main = 0.8))


dev.off()
```



```{r, echo=FALSE, message = FALSE, warning=FALSE}
plot_lag_effect_for_percentile <- function(cpred, percentile, temp_data_matrix, temp_median,
                                           xlab = "Days Relative to Conception (0 = 30 Days Before)",
                                           ylab = "Change in Log Odds of Higher Stress Category",
                                           main = NULL, tag = NULL) {
  # nearest value on cpred grid
  temp_value <- round(quantile(as.vector(temp_data_matrix), percentile/100, na.rm=TRUE), 1)
  row_vals <- as.numeric(rownames(cpred$matfit))
  idx <- which.min(abs(row_vals - temp_value))
  used_val <- row_vals[idx]

  matfit <- cpred$matfit[idx, ]; matlow <- cpred$matlow[idx, ]; mathigh <- cpred$mathigh[idx, ]
  names(matfit) <- names(matlow) <- names(mathigh) <- sub("^lag", "", names(matfit))
  lags <- as.numeric(names(matfit))

  if (is.null(main)) {
    main <- paste0("Effect of ", used_val, "°C (", percentile, "th) vs median (", temp_median, "°C)")
  }

  plot(lags, matfit, type="l", lwd=2,
       ylim=range(c(matlow, mathigh), na.rm=TRUE),
       xlab=xlab, ylab=ylab, main=main)
  polygon(c(lags, rev(lags)), c(mathigh, rev(matlow)),
          col=rgb(0,0,0,0.2), border=NA)
  abline(h=0, lty=2)

  if (!is.null(tag)) mtext(tag, side=3, adj=0, line=0.2, font=2, cex=1.1)

  invisible(invisible(NULL))
}


plot_lag_effect_for_percentile(cpred,   1,  temp_data_matrix,   temp_median, main="A: Effect of 22.4°C (1st) vs median (27.4°C)")



# Plot for 1st, 25th, 75th, and 99th percentiles
plot_lag_effect_for_percentile(cpred, 1, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 25, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 75, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 99, temp_data_matrix, temp_median)
```




```{r}
# helper: add ordinal suffix (1 -> 1st, 2 -> 2nd, etc.)
ordinal <- function(n){
  if (n %% 100L %in% c(11L,12L,13L)) return(paste0(n,"th"))
  c("th","st","nd","rd","th")[pmin(5, (n %% 10L)+1)] |>
    (\(suf) paste0(n, suf))()
}

# build a slice dataframe for a given percentile
slice_df <- function(cpred, percentile, exposure_matrix, center_val){
  
  # find temp values to slice at
  at_val  <- as.numeric(quantile(as.vector(exposure_matrix), percentile/100, na.rm = TRUE))
  rows    <- as.numeric(rownames(cpred$matfit))
  idx     <- which.min(abs(rows - at_val))
  used    <- rows[idx]
  
  # extract predictions
  fit     <- cpred$matfit[idx, ] 
  low <- cpred$matlow[idx, ] 
  high <- cpred$mathigh[idx, ]
  
  # extract lags 
  lags    <- if (all(grepl("^lag", names(fit)))) {
    as.numeric(sub("^lag","",names(fit))) 
    } else {
      seq_along(fit)-1
    }
  
  # build tibble + exponentiate for odds ratios
  tibble(lag = lags,
         effect = as.numeric(fit),
         ci_low = as.numeric(low),
         ci_high = as.numeric(high),
         OR = exp(fit),
         OR_low  = exp(low),
         OR_high = exp(high),
         used_val = used,
         center_val = center_val,
         pct = percentile)
}

library(ggplot2)
library(patchwork)
library(grid)

# one panel with in-plot annotations
make_slice_panel_or <- function(df, letter){

  ann_top <- letter
  ann_sub <- sprintf("%.1f°C (%s) vs median (%.1f°C)",
                     df$used_val[1], ordinal(df$pct[1]), df$center_val[1])

  ggplot(df, aes(lag, OR)) +
    geom_ribbon(aes(ymin = OR_low, ymax = OR_high), alpha = 0.20) +
    geom_line(size = 0.8) +
    geom_hline(yintercept = 1, linetype = 2) +
    annotate("text", x = -Inf, y = Inf, label = ann_top,
             hjust = -0.5, vjust = 0.75, fontface = "bold", size = 5) +
    annotate("text", x = -Inf, y = Inf, label = ann_sub,
             hjust = -0.5, vjust = 2.4, size = 3.5) +
    labs(x = "Days Relative to Conception (0 = 30 Days Before)",
         y = "Odds Ratio (higher stress category)") +
    scale_x_continuous(
      limits = c(0, 250),
      breaks = c(0, 30, 60, 90, 120, 150, 180, 210, 240),
      labels = c("-30", "0", "30", "60", "90", "120", "150", "180", "210")) +
    geom_vline(xintercept = 30, linetype = 3, color = "black", linewidth = 0.4) +
    coord_cartesian(clip = "off") +
    theme_classic(base_size = 12) +
    theme(
      plot.margin = margin(10, 12, 6, 12),
      plot.title = element_blank(),
      axis.text = element_text(size = 11.5, color = "black"),
      axis.title = element_text(size = 12.5, color = "black")
    ) 
}


## ---------- Build the four MAX (adjusted for Δ) panels ----------
A <- make_slice_panel_or(slice_df(cpred, 1, temp_data_matrix, temp_median), "A") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

B <- make_slice_panel_or(slice_df(cpred, 25, temp_data_matrix, temp_median), "B")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

C <- make_slice_panel_or(slice_df(cpred, 75, temp_data_matrix, temp_median), "C")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

D <- make_slice_panel_or(slice_df(cpred, 99, temp_data_matrix, temp_median), "D")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

p_max <- (A + B) / (C + D) +
  plot_layout(guides = "collect", heights = c(1, 1)) &
  theme(plot.margin = margin(t = 5, r = 5, b = 10, l = 10))  # ← increase b (bottom)

# Save to high-resolution PNG
# png(here("outputs", "Figure1_Maxtw.png"), width = 9, height = 6, units = "in", res = 600)

pdf(here("outputs", "Figure_S10.pdf"),
    width = 8, height = 6)


# Draw your combined figure
print(p_max)

# Add shared x/y labels
grid::grid.text("Days relative to conception (0 = conception)",
                y = unit(0.015, "npc"), gp = grid::gpar(fontsize = 12))
grid::grid.text("Odds ratio (higher stress category)",
                rot = 90, x = unit(0.015, "npc"),
                gp = grid::gpar(fontsize = 12.5))

# Close the graphics device — this actually writes the file
dev.off()

```


```{r}
## pull out specific day windows where there is significance to mention in paper
slice_25 <- slice_df(cpred, 25, temp_data_matrix, temp_median)

sig_25 <- slice_25 %>%
  filter(OR_low > 1 | OR_high < 1)

sig_25

slice_99 <- slice_df(cpred, 99, temp_data_matrix, temp_median)

sig_99 <- slice_99 %>%
  filter(OR_low > 1 | OR_high < 1)

sig_99
```


```{r}

temp_w_season <- temp_lags_long %>%
  mutate(
    month  = month(date),
    season = case_when(
      month %in% 4:10 ~ "wet",
      month %in% c(1,2,3,11,12) ~ "dry",
      TRUE ~ NA_character_
    ),
    season = factor(season, levels = c("dry", "wet"))
  )

make_density_df <- function(df, group_vars = c("mstudyid")) {
  df %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(dens = list(density(temp_metric, na.rm = TRUE)), .groups = "drop") %>%
    mutate(x = map(dens, "x"),
           y = map(dens, "y")) %>%
    unnest(c(x, y))
}


# overall per-person
density_overall <- make_density_df(temp_w_season, group_vars = c("mstudyid"))

# dry per-person
density_dry <- make_density_df(filter(temp_w_season, season == "dry"),
                               group_vars = c("mstudyid"))

# wet per-person
density_wet <- make_density_df(filter(temp_w_season, season == "wet"),
                               group_vars = c("mstudyid"))

# pop (overall)
pop_overall <- {
  d <- density(temp_w_season$temp_metric, na.rm = TRUE)
  tibble(x = d$x, y = d$y)
}

# pop (dry / wet)
pop_dry <- {
  d <- density(filter(temp_w_season, season == "dry")$temp_metric, na.rm = TRUE)
  tibble(x = d$x, y = d$y)
}

pop_wet <- {
  d <- density(filter(temp_w_season, season == "wet")$temp_metric, na.rm = TRUE)
  tibble(x = d$x, y = d$y)
}

# percentiles (overall to match DLNM)
qs <- quantile(
  temp_w_season$temp_metric,
  probs = c(0.01, 0.25, 0.5, 0.75, 0.99),
  na.rm = TRUE
)

# shared x-limits so panels are comparable
xlims <- range(temp_w_season$temp_metric, na.rm = TRUE)


make_panel <- function(density_df, pop_df, x_label = NULL) {
  ggplot(density_df, aes(x = x, y = y, group = mstudyid)) +
    geom_line(color = "black", alpha = 0.05, linewidth = 0.4) +
    geom_line(data = pop_df, aes(x = x, y = y),
              inherit.aes = FALSE,
              color = "black", linewidth = 1.1) +
    geom_vline(
      xintercept = qs[c("1%", "25%", "75%", "99%")],
      color = "red3",
      linetype = "dashed",
      linewidth = 0.6
    ) +
    geom_vline(
      xintercept = qs["50%"],
      color = "red",
      linetype = "solid",
      linewidth = 0.9
    ) +
    geom_hline(yintercept = 0, color = "grey80", linewidth = 0.4) +
    coord_cartesian(xlim = xlims) +
    labs(x = x_label, y = "Density") +
    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      axis.line  = element_line(linewidth = 0.9, color = "black"),
      axis.ticks = element_line(linewidth = 0.6, color = "black")
    )
}


p_overall <- make_panel(density_overall, pop_overall, x_label = NULL) +
  ggtitle("A. Full exposure window")

p_dry <- make_panel(density_dry, pop_dry, x_label = NULL) +
  ggtitle("B. Dry season")

p_wet <- make_panel(density_wet, pop_wet, x_label = "Daily Maximum TW (°C)") +
  ggtitle("C. Wet season")

library(patchwork)

full_temp_dist_plot <- (p_overall / p_dry / p_wet) +
  plot_layout(heights = c(1, 1, 1))


ggsave(here("outputs", "figureS3_tw_density_plot.png"),
       full_temp_dist_plot, 
       width = 6, 
       height = 9, 
       dpi = 400)
```



#### Financial Negative Event 

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# set up DLNM model parameters, fit model, and get predictions

## SET UP PARAMS

# predictor DLNM model parameters -- captures the relationship between the predictor variable (temp) and the outcome
var_arg = list(fun = "ns", df = 4) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# lag DLNM model parameters -- captures the lagged effects of the predictor variable (temp) over time.
lag_arg = list(fun = "ns", df = 3) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# cross-basis matrix for DLNM
cb = crossbasis(temp_data_matrix, lag = c(0, ncol(temp_data_matrix) - 1), argvar = var_arg, arglag = lag_arg) # basis for the predictor and lag 

# ordered logistic regression model
model_polr <- polr(
  formula = neg_fin_groups ~ cb + age + bmi + wealthindex + season_delivery + pregchn + married_bin + medlev_bin,
  data = temperature_lags_wide,
  method = "logistic" 
)


# GET PREDICTIONS (in reference to median temp)
temp_median <- round(median(as.vector(temp_data_matrix)), 1) # centering value (median). 
crysis_temp_median <- temp_median

# generate predictions for a DLNM, set centering value as the median of all temperatures
cpred = crosspred(cb, 
                  model_polr, 
                  ci.level = 0.95, # 95% confidence intervals calculated around predicted values
                  cen = temp_median, # predictions are relative to this baseline temperature.
                  by = 0.1) # predictions are generated at intervals of 0.1 units of temperature


# plot contour map of 3d plot
plot(
  cpred, 
  "contour", 
  main = paste0("Effects of Temperature and Lag on Log Odds \nof Higher Financial Stress Category Relative \nto Median Max Temperature (",temp_median, "°C)"), 
  xlab = "Max Temperature (°C)", 
  ylab = "Days Relative to Conception (0 = 30 Days Before)", 
  key.title = title(main = "Change in \nLog Odds", cex.main = 0.8)
)
```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# Plot for 1st, 25th, 75th, and 99th percentiles
plot_lag_effect_for_percentile(cpred, 1, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 25, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 75, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 99, temp_data_matrix, temp_median)
```


```{r}

## ---------- Build the four MAX (adjusted for Δ) panels ----------
A <- make_slice_panel_or(slice_df(cpred, 1, temp_data_matrix, temp_median), "A") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

B <- make_slice_panel_or(slice_df(cpred, 25, temp_data_matrix, temp_median), "B")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

C <- make_slice_panel_or(slice_df(cpred, 75, temp_data_matrix, temp_median), "C")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

D <- make_slice_panel_or(slice_df(cpred, 99, temp_data_matrix, temp_median), "D")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

p_max <- (A + B) / (C + D) +
  plot_layout(guides = "collect", heights = c(1, 1)) &
  theme(plot.margin = margin(t = 5, r = 5, b = 10, l = 10))  # ← increase b (bottom)

# Save to high-resolution PNG
png(here("outputs", "FigureS8_financial_Maxtw.png"), width = 9, height = 6, units = "in", res = 600)

# Draw your combined figure
print(p_max)

# Add shared x/y labels
grid::grid.text("Days relative to conception (0 = conception)",
                y = unit(0.015, "npc"), gp = grid::gpar(fontsize = 12))
grid::grid.text("Odds ratio (higher stress category)",
                rot = 90, x = unit(0.01, "npc"), gp = grid::gpar(fontsize = 12))

# Close the graphics device — this actually writes the file
dev.off()
```

#### Relationship Negative Event 

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# set up DLNM model parameters, fit model, and get predictions

## SET UP PARAMS

# predictor DLNM model parameters -- captures the relationship between the predictor variable (temp) and the outcome
var_arg = list(fun = "ns", df = 4) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# lag DLNM model parameters -- captures the lagged effects of the predictor variable (temp) over time.
lag_arg = list(fun = "ns", df = 3) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# cross-basis matrix for DLNM
cb = crossbasis(temp_data_matrix, lag = c(0, ncol(temp_data_matrix) - 1), argvar = var_arg, arglag = lag_arg) # basis for the predictor and lag 

# ordered logistic regression model
model_polr <- polr(
  formula = neg_rel_groups ~ cb + age + bmi + wealthindex + season_delivery + pregchn + married_bin + medlev_bin,
  data = temperature_lags_wide,
  method = "logistic" 
)


# GET PREDICTIONS (in reference to median temp)
temp_median <- round(median(as.vector(temp_data_matrix)), 1) # centering value (median). 
crysis_temp_median <- temp_median

# generate predictions for a DLNM, set centering value as the median of all temperatures
cpred = crosspred(cb, 
                  model_polr, 
                  ci.level = 0.95, # 95% confidence intervals calculated around predicted values
                  cen = temp_median, # predictions are relative to this baseline temperature.
                  by = 0.1) # predictions are generated at intervals of 0.1 units of temperature


# plot contour map of 3d plot
plot(
  cpred, 
  "contour", 
  main = paste0("Effects of Temperature and Lag on Log Odds \nof Higher Relationship Stress Category Relative \nto Median Max Temperature (",temp_median, "°C)"), 
  xlab = "Max Temperature (°C)", 
  ylab = "Days Relative to Conception (0 = 30 Days Before)", 
  key.title = title(main = "Change in \nLog Odds", cex.main = 0.8)
)
```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# Plot for 1st, 25th, 75th, and 99th percentiles

plot_lag_effect_for_percentile(cpred, 1, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 25, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 75, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 99, temp_data_matrix, temp_median)
```


```{r}
slice_99 <- slice_df(cpred, 99, temp_data_matrix, temp_median)

sig_99 <- slice_99 %>%
  filter(OR_low > 1 | OR_high < 1)

sig_99
```



#### Home Negative Event 

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# set up DLNM model parameters, fit model, and get predictions

## SET UP PARAMS

# predictor DLNM model parameters -- captures the relationship between the predictor variable (temp) and the outcome
var_arg = list(fun = "ns", df = 4) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# lag DLNM model parameters -- captures the lagged effects of the predictor variable (temp) over time.
lag_arg = list(fun = "ns", df = 3) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# cross-basis matrix for DLNM
cb = crossbasis(temp_data_matrix, lag = c(0, ncol(temp_data_matrix) - 1), argvar = var_arg, arglag = lag_arg) # basis for the predictor and lag 


# ordered logistic regression model
model_glm <- glm(
  neg_home_groups ~ cb + age + bmi + wealthindex + season_delivery + pregchn + married_bin + medlev_bin,
  data   = temperature_lags_wide,
  family = binomial(link = "logit")
)


# GET PREDICTIONS (in reference to median temp)
temp_median <- round(median(as.vector(temp_data_matrix)), 1) # centering value (median). 
crysis_temp_median <- temp_median

# generate predictions for a DLNM, set centering value as the median of all temperatures
cpred <- crosspred(
  cb,
  model_glm,
  ci.level = 0.95,
  cen = temp_median,
  by = 0.1,
)

# plot contour map of 3d plot
plot(
  cpred, 
  "contour", 
  main = paste0("Effects of Temperature and Lag on Odds Ratio \nof Home Stress Relative \nto Median Max Temperature (",temp_median, "°C)"), 
  xlab = "Max Temperature (°C)", 
  ylab = "Days Relative to Conception (0 = 30 Days Before)", 
  key.title = title(main = "Odds ratio \n(vs median)", cex.main = 0.8)
)
```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# --- Helpers (RR scale only) -----------------------------------------------

.get_rr <- function(cpred) {
  if (is.null(cpred$matRRfit) || is.null(cpred$matRRlow) || is.null(cpred$matRRhigh)) {
    stop("cpred does not contain RR matrices. Re-run crosspred() normally (no 'type=' arg).")
  }
  list(fit = cpred$matRRfit, low = cpred$matRRlow, high = cpred$matRRhigh)
}

.closest_row <- function(row_names_num, target) {
  # returns index of closest row by numeric distance
  which.min(abs(row_names_num - target))
}

# --- Main: percentile lag plot on RR scale to match heatmap ----------------

plot_lag_effect_for_percentile_rr <- function(
  cpred, percentile, temp_data_matrix, temp_median,
  xlab = "Days Relative to Conception (0 = 30 Days Before)",
  ylab = "Relative Risk (vs median temperature)",
  highlight_signif = TRUE,  # mark lags where CI excludes 1
  point_cex = 0.7
) {
  rr <- .get_rr(cpred)

  # Temperature for the requested percentile (from your data)
  target_temp <- as.numeric(quantile(as.vector(temp_data_matrix), percentile/100, na.rm = TRUE))

  # Find the closest available prediction row in cpred (rows are temps)
  temps_in_cpred <- as.numeric(rownames(rr$fit))
  if (anyNA(temps_in_cpred)) stop("Row names of cpred RR matrices must be numeric temperatures.")
  idx <- .closest_row(temps_in_cpred, target_temp)
  picked_temp <- temps_in_cpred[idx]

  # Extract RR and CI across lags for that temperature
  rr_fit  <- as.numeric(rr$fit [idx, ])
  rr_low  <- as.numeric(rr$low [idx, ])
  rr_high <- as.numeric(rr$high[idx, ])

  # Lag axis from column names like "lag0", "lag1", ...
  lag_vals <- as.numeric(sub("lag", "", colnames(rr$fit)))

  # y-limits from finite values
  yy_min <- min(rr_low [is.finite(rr_low )], na.rm = TRUE)
  yy_max <- max(rr_high[is.finite(rr_high)], na.rm = TRUE)

  # Plot RR curve + CI ribbon
  plot(lag_vals, rr_fit, type = "l", lwd = 2,
       ylim = c(yy_min, yy_max),
       xlab = xlab, ylab = ylab,
       main = paste0("Effect of ", round(picked_temp, 1), "°C (", percentile,
                     "th percentile) vs median (", round(temp_median, 1), "°C)"))
  polygon(c(lag_vals, rev(lag_vals)),
          c(rr_high,  rev(rr_low)),
          col = rgb(0, 0, 0, 0.2), border = NA)

  # Baseline for RR
  abline(h = 1, lty = 2)

  # highlight lags where CI excludes 1 (statistically significant)
  if (isTRUE(highlight_signif)) {
    sig_hi <- which(rr_low  > 1 & is.finite(rr_low))
    sig_lo <- which(rr_high < 1 & is.finite(rr_high))
    if (length(sig_hi)) points(lag_vals[sig_hi], rr_fit[sig_hi], pch = 16, cex = point_cex)
    if (length(sig_lo)) points(lag_vals[sig_lo], rr_fit[sig_lo], pch = 16, cex = point_cex)
    legend("topright", inset = 0.01, bty = "n",
           legend = c("CI excludes 1"), pch = 16, cex = 0.8)
  }

  invisible(data.frame(
    lag = lag_vals,
    RR = rr_fit,
    ci_low = rr_low,
    ci_high = rr_high,
    temp_value = picked_temp,
    percentile = percentile
  ))
}

# Plot for 1st, 25th, 75th, and 99th percentiles
plot_lag_effect_for_percentile_rr(cpred,  1, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile_rr(cpred, 25, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile_rr(cpred, 75, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile_rr(cpred, 99, temp_data_matrix, temp_median)
```




### Min tw

```{r, echo=FALSE, message = FALSE, warning=FALSE}
## Create Daily Lag Frame:
max_gestation_days <- 210
pre_conception_days <- 30

crisis_dlnm_eligible <- crisis_simple %>% 
  filter(gestage_at_survey >= max_gestation_days) %>% # ~37 weeks
  filter(!is.na(bmi))


# Function to process temperature data for the selected metric
process_temperature_data <- function(era5_data, metric) {
  era5_data %>%
    select(community, date, !!sym(metric)) %>%
    rename(temp_metric = !!sym(metric)) %>%
    distinct()
}

# Process temp data
temp_data_clean <- process_temperature_data(era5_data = era5, metric = "min_tw")



# Generate lags relative to conception
create_conception_aligned_lags <- function(df, temp_df) {
  df %>%
    select(mstudyid, vname, date_conception) %>%
    rowwise() %>%
    mutate(day_relative = list(-pre_conception_days:max_gestation_days),
           date = list(date_conception + day_relative)) %>%
    unnest(c(day_relative, date)) %>%
    left_join(temp_df, by = c("vname" = "community", "date" = "date"))  # join temp
}


temp_lags_long <- create_conception_aligned_lags(crisis_dlnm_eligible, temp_data_clean)

# Step 4: Pivot to wide format (like DLNM expects)
temperature_lags_wide <- temp_lags_long %>%
  select(mstudyid, date_conception, day_relative, temp_metric) %>%
  pivot_wider(
    names_from = day_relative,
    values_from = temp_metric,
    names_prefix = "day_"
  ) %>%
  left_join(crisis_dlnm_eligible %>% select(mstudyid, sum_nds_category, neg_fin_groups, neg_rel_groups, neg_home_groups, age, bmi, wealthindex, season_delivery, pregchn, married_bin, medlev_bin), by = "mstudyid") %>%
  filter(!is.na(day_0))

temp_data_matrix <- temperature_lags_wide %>%
  select(starts_with("day_")) %>%
  as.matrix()



# set up DLNM model parameters, fit model, and get predictions

## SET UP PARAMS

# predictor DLNM model parameters -- captures the relationship between the predictor variable (temp) and the outcome
var_arg = list(fun = "ns", df = 4) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# lag DLNM model parameters -- captures the lagged effects of the predictor variable (temp) over time.
lag_arg = list(fun = "ns", df = 3) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# cross-basis matrix for DLNM
cb = crossbasis(temp_data_matrix, lag = c(0, ncol(temp_data_matrix) - 1), argvar = var_arg, arglag = lag_arg) # basis for the predictor and lag 

# ordered logistic regression model
model_polr <- polr(
  formula = sum_nds_category ~ cb + age + bmi + wealthindex + season_delivery + pregchn + married_bin + medlev_bin,
  data = temperature_lags_wide,
  method = "logistic" 
)



# GET PREDICTIONS (in reference to median temp)
temp_median <- round(median(as.vector(temp_data_matrix)), 1) # centering value (median). 
crysis_temp_median <- temp_median


## check temp dist quickly 
temp_vec <- as.vector(temp_data_matrix)
summary(temp_vec)
quantile(temp_vec, probs = c(0.01, 0.05, 0.1, 0.25, 0.5))
mean(temp_vec < 17, na.rm = TRUE) * 100   # % of observations below 17 °C

ggplot(data.frame(temp_vec), aes(x = temp_vec)) +
  geom_histogram(binwidth = 0.5, fill = "steelblue", color = "black") +
  geom_vline(xintercept = 17, linetype = "dashed", color = "red") +
  labs(x = "Min Temperature (°C)", y = "Frequency",
       title = "Distribution of Minimum Temperature Values") +
  theme_minimal()


# generate predictions for a DLNM, set centering value as the median of all temperatures
cpred = crosspred(cb, 
                  model_polr, 
                  ci.level = 0.95, # 95% confidence intervals calculated around predicted values
                  cen = temp_median, # predictions are relative to this baseline temperature.
                  by = 0.1) # predictions are generated at intervals of 0.1 units of temperature




png(here("outputs", "figure_S2_contour_min_tw.png"), width = 8, height = 6, units = "in", res = 300)

par(mar = c(4.8, 5.6, 2.2, 6.2))  # bottom, left, top, right

plot(cpred, "contour",
     xlab = "Min temperature (°C)",
     ylab = "Days relative to conception (0 = 30 days preconception)",
     key.title = title(main = "Change in \nlog odds", cex.main = 0.8))

dev.off()




# Plot for 1st, 25th, 75th, and 99th percentiles
plot_lag_effect_for_percentile(cpred, 1, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 25, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 75, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 99, temp_data_matrix, temp_median)
```


## MIN tw FIGURE FOR PAPER

```{r}

## ---------- Build the four Min (adjusted for Δ) panels ----------
A <- make_slice_panel_or(slice_df(cpred, 1, temp_data_matrix, temp_median), "A") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

B <- make_slice_panel_or(slice_df(cpred, 25, temp_data_matrix, temp_median), "B")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

C <- make_slice_panel_or(slice_df(cpred, 75, temp_data_matrix, temp_median), "C")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

D <- make_slice_panel_or(slice_df(cpred, 99, temp_data_matrix, temp_median), "D")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

p_max <- (A + B) / (C + D) +
  plot_layout(guides = "collect", heights = c(1, 1)) &
  theme(plot.margin = margin(t = 5, r = 5, b = 10, l = 10))  # ← increase b (bottom)

# Save to high-resolution PNG
#png(here("outputs", "Figure2_Mintw.png"), width = 9, height = 6, units = "in", res = 600)

pdf(here("outputs", "Figure_S13.pdf"),
    width = 8, height = 6)

# Draw your combined figure
print(p_max)

# Add shared x/y labels
grid::grid.text("Days relative to conception (0 = conception)",
                y = unit(0.015, "npc"), gp = grid::gpar(fontsize = 12))
grid::grid.text("Odds ratio (higher stress category)",
                rot = 90, x = unit(0.015, "npc"),
                gp = grid::gpar(fontsize = 12.5))

# Close the graphics device — this actually writes the file
dev.off()
```


```{r}
## pull out specific day windows where there is significance to mention in paper
slice_75 <- slice_df(cpred, 75, temp_data_matrix, temp_median)

sig_75 <- slice_75 %>%
  filter(OR_low > 1 | OR_high < 1)

sig_75




slice_99 <- slice_df(cpred, 99, temp_data_matrix, temp_median)

sig_99 <- slice_99 %>%
  filter(OR_low > 1 | OR_high < 1)

sig_99
```

### tw Delta (max-min temp) to examine effect of cooling

```{r, echo=FALSE, message = FALSE, warning=FALSE}
## Create Daily Lag Frame:
max_gestation_days <- 210
pre_conception_days <- 30

crisis_dlnm_eligible <- crisis_simple %>% 
  filter(gestage_at_survey >= max_gestation_days) %>% # ~37 weeks
  filter(!is.na(bmi))


# Function to process temperature data for the selected metric
process_temperature_data <- function(era5_data, metric) {
  era5_data %>%
    select(community, date, !!sym(metric)) %>%
    rename(temp_metric = !!sym(metric)) %>%
    distinct()
}

# Process temp data
temp_data_clean <- process_temperature_data(era5_data = era5, metric = "tw_delta")



# Generate lags relative to conception
create_conception_aligned_lags <- function(df, temp_df) {
  df %>%
    select(mstudyid, vname, date_conception) %>%
    rowwise() %>%
    mutate(day_relative = list(-pre_conception_days:max_gestation_days),
           date = list(date_conception + day_relative)) %>%
    unnest(c(day_relative, date)) %>%
    left_join(temp_df, by = c("vname" = "community", "date" = "date"))  # join temp
}


temp_lags_long <- create_conception_aligned_lags(crisis_dlnm_eligible, temp_data_clean)

# Step 4: Pivot to wide format (like DLNM expects)
temperature_lags_wide <- temp_lags_long %>%
  select(mstudyid, date_conception, day_relative, temp_metric) %>%
  pivot_wider(
    names_from = day_relative,
    values_from = temp_metric,
    names_prefix = "day_"
  ) %>%
  left_join(crisis_dlnm_eligible %>% select(mstudyid, sum_nds_category, neg_fin_groups, neg_rel_groups, neg_home_groups, age, bmi, wealthindex, pregchn, season_delivery, married_bin, medlev_bin), by = "mstudyid") %>%
  filter(!is.na(day_0))

temp_data_matrix <- temperature_lags_wide %>%
  select(starts_with("day_")) %>%
  as.matrix()

```

```{r, echo=FALSE, message = FALSE, warning=FALSE}
# set up DLNM model parameters, fit model, and get predictions

## SET UP PARAMS

# predictor DLNM model parameters -- captures the relationship between the predictor variable (temp) and the outcome
var_arg = list(fun = "ns", df = 4) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# lag DLNM model parameters -- captures the lagged effects of the predictor variable (temp) over time.
lag_arg = list(fun = "ns", df = 3) # smooth cubic polynomial (degree = 3) between the boundary knots (no internal knots)

# cross-basis matrix for DLNM
cb = crossbasis(temp_data_matrix, lag = c(0, ncol(temp_data_matrix) - 1), argvar = var_arg, arglag = lag_arg) # basis for the predictor and lag 

# ordered logistic regression model
model_polr <- polr(
  formula = sum_nds_category ~ cb + age + bmi + wealthindex + season_delivery + pregchn + married_bin + medlev_bin,
  data = temperature_lags_wide,
  method = "logistic" 
)


# GET PREDICTIONS (in reference to median temp)
temp_median <- round(median(as.vector(temp_data_matrix)), 1) # centering value (median). 
crysis_temp_median <- temp_median

# generate predictions for a DLNM, set centering value as the median of all temperatures
cpred = crosspred(cb, 
                  model_polr, 
                  ci.level = 0.95, # 95% confidence intervals calculated around predicted values
                  cen = temp_median, # predictions are relative to this baseline temperature.
                  by = 0.1) # predictions are generated at intervals of 0.1 units of temperature


# plot contour map of 3d plot
plot(
  cpred, 
  "contour", 
  main = paste0("Effects of Temperature Delta and Lag on Log Odds \nof Higher Stress Category Relative \nto Median Delta Temperature (",temp_median, "°C)"), 
  xlab = "Temperature Delta (°C)", 
  ylab = "Days Relative to Conception (0 = 30 Days Before)", 
  key.title = title(main = "Change in \nLog Odds", cex.main = 0.8)
)
```



```{r, echo=FALSE, message = FALSE, warning=FALSE}
plot_lag_effect_for_percentile <- function(cpred, percentile, temp_data_matrix, temp_median,
                                           xlab = "Days Relative to Conception (0 = 30 Days Before)",
                                           ylab = "Change in Log Odds of Higher Stress Category") {
  
  # Get temperature at desired percentile
  temp_value <- round(quantile(as.vector(temp_data_matrix), percentile / 100), 1)
  
  # Extract effect estimates and CI for that temperature
  matfit <- cpred$matfit[rownames(cpred$matfit) == temp_value, ]
  matlow <- cpred$matlow[rownames(cpred$matlow) == temp_value, ]
  mathigh <- cpred$mathigh[rownames(cpred$mathigh) == temp_value, ]
  
  # Clean lag names
  names(matfit) <- names(matlow) <- names(mathigh) <- sub("lag", "", names(matfit))
  
  # Create plot
  plot(
    x = as.numeric(names(matfit)),
    y = matfit,
    type = "l",
    lwd = 2,
    ylim = c(min(matlow), max(mathigh)),
    xlab = xlab,
    ylab = ylab,
    main = paste0("Effect of ", temp_value, "\u00B0C (", percentile, "th Percentile) \nvs Median (", temp_median, "\u00B0C)")
  )
  
  # Add shaded confidence interval
  polygon(
    x = c(as.numeric(names(mathigh)), rev(as.numeric(names(matlow)))),
    y = c(mathigh, rev(matlow)),
    col = rgb(0, 0, 0, alpha = 0.2),
    border = NA
  )
  
  abline(h = 0, lty = 2)
  
  # Return data frame if needed
  return(invisible(data.frame(
    lag = as.numeric(names(matfit)),
    effect = matfit,
    ci_low = matlow,
    ci_high = mathigh,
    temp_value = temp_value,
    percentile = percentile
  )))
}


# Plot for 1st, 25th, 75th, and 99th percentiles
plot_lag_effect_for_percentile(cpred, 1, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 25, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 75, temp_data_matrix, temp_median)
plot_lag_effect_for_percentile(cpred, 99, temp_data_matrix, temp_median)
```






### tw MAX + DELTA INTERACTION


```{r, echo=FALSE, message = FALSE, warning=FALSE}
## --- Setup shared calendar & pull both metrics ---
max_gestation_days <- 210
pre_conception_days <- 30

crisis_dlnm_eligible <- crisis_simple %>% 
  filter(gestage_at_survey >= max_gestation_days) %>% # ~37 weeks
  filter(!is.na(bmi))


# a) conception-aligned calendar (one row per subject-day)
make_calendar <- function(df) {
  df %>%
    select(mstudyid, vname, date_conception) %>%
    rowwise() %>%
    mutate(day_relative = list(-pre_conception_days:max_gestation_days),
           date = list(date_conception + day_relative)) %>%
    unnest(c(day_relative, date))
}

cal_long <- make_calendar(crisis_dlnm_eligible)

# b) join BOTH exposures to the same long frame
expos_long <- cal_long %>%
  left_join(
    era5 %>%
      select(community, date, max_tw, min_tw, tw_delta),
    by = c("vname" = "community", "date" = "date")
  )

## --- Pivot to wide twice: one matrix for max, one for delta ---
wide_max <- expos_long %>%
  select(mstudyid, date_conception, day_relative, max_tw) %>%
  pivot_wider(names_from = day_relative, values_from = max_tw, names_prefix = "max_")

wide_delta <- expos_long %>%
  select(mstudyid, date_conception, day_relative, tw_delta) %>%
  pivot_wider(names_from = day_relative, values_from = tw_delta, names_prefix = "delta_")

# c) align rows and add covariates
design <- crisis_dlnm_eligible %>%
  select(mstudyid, sum_nds_category, age, bmi, wealthindex, pregchn, season_delivery, married_bin, medlev_bin) %>%
  left_join(wide_max,   by = c("mstudyid")) %>%
  left_join(wide_delta, by = c("mstudyid")) %>%
  # require exposure present at "day_0" for BOTH metrics
  filter(!is.na(max_0), !is.na(delta_0))

# d) build matrices
mat_max   <- design %>% select(starts_with("max_"))   %>% as.matrix()
mat_delta <- design %>% select(starts_with("delta_")) %>% as.matrix()

## --- Cross-bases and model (joint main effects) ---
library(dlnm)
var_arg <- list(fun = "ns", df = 4)
lag_arg <- list(fun = "ns", df = 3)

cb_max   <- crossbasis(mat_max,   lag = c(0, ncol(mat_max)   - 1), argvar = var_arg, arglag = lag_arg)
cb_delta <- crossbasis(mat_delta, lag = c(0, ncol(mat_delta) - 1), argvar = var_arg, arglag = lag_arg)

# Fit ordered logistic with BOTH cross-bases
model_joint <- MASS::polr(
  sum_nds_category ~ cb_max + cb_delta + age + bmi + wealthindex + season_delivery + pregchn + married_bin + medlev_bin,
  data = design,
  method = "logistic"
)

## --- Predictions: adjusted surfaces ---
cen_max   <- round(median(as.vector(mat_max),   na.rm = TRUE), 1)
cen_delta <- round(median(as.vector(mat_delta), na.rm = TRUE), 1)

pred_max_adj   <- crosspred(cb_max,   model_joint, cen = cen_max,   by = 0.1)
pred_delta_adj <- crosspred(cb_delta, model_joint, cen = cen_delta, by = 0.1)

# Example plot (max tw adjusted for delta)
plot(pred_max_adj, "contour",
     main = paste0("Max tw × Lag (adj for Δ), centered at ", cen_max, "°C"),
     xlab = "Max tw (°C)",
     ylab = "Days Relative to Conception (0 = 30 Days Before)",
     key.title = title(main = "Δ Log-Odds", cex.main = 0.8))

# Example plot (delta adjusted for max)
plot(pred_delta_adj, "contour",
     main = paste0("Overnight Cooling (Δ) × Lag (adj for Max), centered at ", cen_delta, "°C"),
     xlab = "tw Δ (°C)",
     ylab = "Days Relative to Conception (0 = 30 Days Before)",
     key.title = title(main = "Δ Log-Odds", cex.main = 0.8))

```




```{r, echo=FALSE, message = FALSE, warning=FALSE}
slice_plot <- function(cpred,
                       exposure_matrix,
                       center_val,
                       percentile = NULL,   # e.g., 1, 25, 75, 99
                       at_val     = NULL,   # or a fixed value like 31.0
                       adj_for    = "Δ",    # label: what this is adjusted for
                       exposure_label = "", # optional label to prepend in title
                       xlab = "Days Relative to Conception (0 = 30 Days Before)",
                       ylab = "Δ Log-Odds (higher stress category)",
                       main = NULL) {

  if (is.null(at_val)) {
    stopifnot(!is.null(percentile))
    at_val <- as.numeric(quantile(as.vector(exposure_matrix),
                                  probs = percentile/100, na.rm = TRUE))
  }

  # find nearest grid value on cpred
  row_vals <- as.numeric(rownames(cpred$matfit))
  idx <- which.min(abs(row_vals - at_val))
  used_val <- row_vals[idx]

  matfit  <- cpred$matfit[idx, ]
  matlow  <- cpred$matlow[idx, ]
  mathigh <- cpred$mathigh[idx, ]

  # lag axis
  lag_names <- names(matfit)
  lags <- if (all(grepl("^lag", lag_names))) as.numeric(sub("^lag", "", lag_names)) else seq_along(matfit) - 1

  # auto title if none supplied
  if (is.null(main)) {
    ptxt <- if (!is.null(percentile)) paste0(" (", percentile, "th percentile)") else ""
    prefix <- if (nzchar(exposure_label)) paste0(exposure_label, " ") else ""
    main <- sprintf("Effect at %s%.1f°C%s compared to median (%.1f°C), adj for %s",
                    prefix, used_val, ptxt, center_val, adj_for)
  }

  # plot
  plot(lags, matfit, type = "l", lwd = 2,
       ylim = range(c(matlow, mathigh), na.rm = TRUE),
       xlab = xlab, ylab = ylab, main = main)
  polygon(c(lags, rev(lags)), c(mathigh, rev(matlow)),
          col = rgb(0, 0, 0, 0.2), border = NA)
  abline(h = 0, lty = 2)

  invisible(tibble(lag = lags, effect = matfit, ci_low = matlow, ci_high = mathigh,
                   requested_val = at_val, used_val = used_val, center_val = center_val))
}



## ---------- CALLS FOR YOUR TWO ADJUSTED SURFACES ----------
## MAX tw slices (adjusted for Δ)
slice_plot(pred_max_adj, mat_max, center_val = cen_max,
           percentile = 1,  adj_for = "Δ",        exposure_label = "Max tw")
slice_plot(pred_max_adj, mat_max, center_val = cen_max,
           percentile = 25, adj_for = "Δ",        exposure_label = "Max tw")
slice_plot(pred_max_adj, mat_max, center_val = cen_max,
           percentile = 75, adj_for = "Δ",        exposure_label = "Max tw")
slice_plot(pred_max_adj, mat_max, center_val = cen_max,
           percentile = 99, adj_for = "Δ",        exposure_label = "Max tw")

## Δ (overnight cooling) slices (adjusted for max tw)
slice_plot(pred_delta_adj, mat_delta, center_val = cen_delta,
           percentile = 1,  adj_for = "max tw", exposure_label = "Δtw")
slice_plot(pred_delta_adj, mat_delta, center_val = cen_delta,
           percentile = 25, adj_for = "max tw", exposure_label = "Δtw")
slice_plot(pred_delta_adj, mat_delta, center_val = cen_delta,
           percentile = 75, adj_for = "max tw", exposure_label = "Δtw")
slice_plot(pred_delta_adj, mat_delta, center_val = cen_delta,
           percentile = 99, adj_for = "max tw", exposure_label = "Δtw")

```




```{r}

## ---------- Build the four MAX (adjusted for Δ) panels ----------
A <- make_slice_panel_or(slice_df(pred_max_adj, 1, mat_max, cen_max), "A") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

B <- make_slice_panel_or(slice_df(pred_max_adj, 25, mat_max, cen_max), "B")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

C <- make_slice_panel_or(slice_df(pred_max_adj, 75, mat_max, cen_max), "C")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

D <- make_slice_panel_or(slice_df(pred_max_adj, 99, mat_max, cen_max), "D")+ 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

p_max <- (A + B) / (C + D) +
  plot_layout(guides = "collect", heights = c(1, 1)) &
  theme(plot.margin = margin(t = 5, r = 5, b = 10, l = 10))  # ← increase b (bottom)

# Save to high-resolution PNG
png(here("outputs", "FigureS7_adjusted_Maxtw.png"), width = 9, height = 6, units = "in", res = 600)

# Draw your combined figure
print(p_max)

# Add shared x/y labels
grid::grid.text("Days Relative to Conception (0 = 30 Days Before)",
                y = unit(0.015, "npc"), gp = grid::gpar(fontsize = 12))
grid::grid.text("Δ Log-Odds (higher stress category)",
                rot = 90, x = unit(0.01, "npc"), gp = grid::gpar(fontsize = 12))

# Close the graphics device — this actually writes the file
dev.off()
```







### Trimester-level ordered logistic regressions

```{r, echo=FALSE, message = FALSE, warning=FALSE}
### Make sure to use crysis term data
crisis_dlnm_eligible <- crisis_dlnm_eligible %>%
  mutate(sum_nds_category = ordered(sum_nds_category, levels = c("low","moderate","high")))


calc_custom_temp_avg <- function(stress_data, temp_data, metric) {
  stress_data %>%
    rowwise() %>%
    mutate(
      # Temperature on the day of the survey
      avg_temp_day_of_survey = temp_data %>%
        filter(community == vname, date == survey_date) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp),
      
      # Temperature the day before the survey
      avg_temp_day_before_survey = temp_data %>%
        filter(community == vname, date == survey_date - 1) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp),
      
      # Average temperature for the week before the survey
      avg_temp_week_before_survey = temp_data %>%
        filter(community == vname, date >= survey_date - 7 & date < survey_date) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp),
      
       # Average temperature for the month before the survey
      avg_temp_month_before_survey = temp_data %>%
        filter(community == vname, date >= survey_date - 30 & date < survey_date) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp),
      
      # Average temperature for the 3 months before the survey
      avg_temp_3_months_before = temp_data %>%
        filter(community == vname, date >= survey_date - 90 & date < survey_date) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp),
      
      # Average temperature for the 6 months before the survey
      avg_temp_6_months_before = temp_data %>%
        filter(community == vname, date >= survey_date - 180 & date < survey_date) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp),
      
      ## TRIMESTER TEMP CALCULATIONS
      
      # Pre-conception temperature average (30 days before conception)
      avg_temp_pre_conception = temp_data %>%
        filter(
          community == vname,
          date >= date_conception - 30 & date < date_conception
        ) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp),
      
      # First trimester temperature average
      avg_temp_first_trimester = temp_data %>%
        filter(
          community == vname,
          date >= date_conception & date < date_conception + 98
        ) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp),
      
      # Second trimester temperature average (require ≥30 days of data)
      avg_temp_second_trimester = {
        x <- temp_data %>%
          filter(community == vname,
                 date >= date_conception + 98,
                 date < date_conception + 196)
        if (nrow(x) >= 30) mean(x[[metric]], na.rm = TRUE) else NA_real_
      },
      
      # Third trimester temperature average
      avg_temp_third_trimester = if (survey_date < datdeliv) {
        # surveyed before delivery → exposures up to survey
        x <- temp_data %>%
          filter(community == vname,
                 date >= date_conception + 196,
                 date <= survey_date)
        if (nrow(x) >= 30) mean(x[[metric]], na.rm = TRUE) else NA_real_
      } else {
        # surveyed after delivery → full trimester up to delivery
        x <- temp_data %>%
          filter(community == vname,
                 date >= date_conception + 196,
                 date <= datdeliv)
        if (nrow(x) >= 30) mean(x[[metric]], na.rm = TRUE) else NA_real_
      }, 
      # Average temperature across the entire pregnancy
      avg_temp_entire_pregnancy = temp_data %>%
        filter(
          community == vname,
          date >= date_conception & date <= survey_date
        ) %>%
        summarise(mean_temp = mean(!!sym(metric), na.rm = TRUE)) %>%
        pull(mean_temp)
    ) %>%
    ungroup()
}



metric <- "max_tw"

stress_data_with_temp <- calc_custom_temp_avg(crisis_dlnm_eligible, era5, metric)

vars <- c(
  "avg_temp_pre_conception",
  "avg_temp_first_trimester",
  "avg_temp_second_trimester",
  "avg_temp_third_trimester",
  "avg_temp_entire_pregnancy"
)

stress_data_with_temp %>%
  select(all_of(vars)) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 30) +
  facet_wrap(~name, scales = "free_x")




##  Fit ordered logits (unadjusted now; add covars later)
models <- list(
  "Pre-Conception"    = polr(sum_nds_category ~ avg_temp_pre_conception + age + bmi + wealthindex + pregchn + season_delivery + married_bin + medlev_bin, data = stress_data_with_temp, method = "logistic", na.action = na.omit),
  
  "First Trimester"   = polr(sum_nds_category ~ avg_temp_first_trimester + age + bmi + wealthindex + pregchn + season_delivery + married_bin + medlev_bin, data = stress_data_with_temp, method = "logistic", na.action = na.omit),
  
  "Second Trimester"  = polr(sum_nds_category ~ avg_temp_second_trimester + age + bmi + wealthindex + pregchn + season_delivery + married_bin + medlev_bin,  data = stress_data_with_temp, method = "logistic", na.action = na.omit),
  
  "Third Trimester"   = polr(sum_nds_category ~ avg_temp_third_trimester + age + bmi + wealthindex + pregchn + season_delivery + married_bin + medlev_bin, data = stress_data_with_temp, method = "logistic", na.action = na.omit),
  
  "Entire Pregnancy"  = polr(sum_nds_category ~ avg_temp_entire_pregnancy + age + bmi + wealthindex + pregchn + season_delivery + married_bin + medlev_bin, data = stress_data_with_temp, method = "logistic", na.action = na.omit)
)


## 4) Helper to tidy polr and drop cut-points
tidy_polr <- function(fit) {
  td <- broom::tidy(fit) %>%
    filter(!str_detect(term, "\\|")) %>%       # drop thresholds
    mutate(p.value = 2*pnorm(abs(statistic), lower.tail = FALSE))
  td
}

## 5) Build MAIN table (exposure effect only, plus CI, P, N)
window_map <- tibble::tibble(
  Model = names(models),
  term  = c("avg_temp_pre_conception","avg_temp_first_trimester",
            "avg_temp_second_trimester","avg_temp_third_trimester",
            "avg_temp_entire_pregnancy")
)

main_tbl <- purrr::map2_dfr(models, window_map$term, ~{
  td <- tidy_polr(.x) %>% filter(term == .y)
  td %>%
  transmute(
    Model = window_map$Model[which(window_map$term == .y)],
    OR = exp(estimate),
    lo = exp(estimate - 1.96*std.error),
    hi = exp(estimate + 1.96*std.error),
    `OR (95% CI)` = sprintf("%.2f (%.2f, %.2f)", OR, lo, hi),
    `P-value` = p.value,
    N = nrow(model.frame(.x))
  )
})

main_table_save <- main_tbl %>%
  select(Model, `OR (95% CI)`, `P-value`, N) %>%
  gt() %>%
  fmt_number(columns = c(`P-value`), decimals = 3) %>%
  cols_align("center")

## 6) Build SUPPLEMENTAL table (full coefficients, per window)

format_p <- function(p) {
  ifelse(p < 0.001, "<0.001",
         ifelse(p < 0.01,
                sprintf("%.3f", p),
                sprintf("%.2f", p)))
}

# Build SUPPLEMENTAL table: OR + CI + p-value
supp_tbl_clean <- purrr::imap_dfr(models, function(m, nm) {
  tidy_polr(m) %>%
    mutate(
      Model = nm,
      OR  = exp(estimate),
      lo  = exp(estimate - 1.96 * std.error),
      hi  = exp(estimate + 1.96 * std.error),
      OR_CI = sprintf("%.2f (%.2f, %.2f)", OR, lo, hi),
      p_formatted = format_p(p.value)
    ) %>%
    select(Model, term, OR_CI, p_formatted)
})

# Convert to gt table
supp_table_save <- supp_tbl_clean %>%
  rename(
    Predictor = term,
    `OR (95% CI)` = OR_CI,
    `P-value` = p_formatted
  ) %>%
  gt(groupname_col = "Model") %>%
  cols_align("center") %>%
  tab_options(
    table.font.size = px(12),
    data_row.padding = px(4),
    heading.align = "left"
  )


# supp_tbl <- purrr::imap_dfr(models, function(m, nm) {
#   tidy_polr(m) %>%
#     mutate(
#       Model = nm,
#       OddsRatio = exp(estimate)
#     ) %>%
#     select(Model, term, estimate, std.error, OddsRatio, p.value)
# })
# 
# supp_table_save <- supp_tbl %>%
#   rename(
#     Predictor = term,
#     Coefficient = estimate,
#     `Std. Error` = std.error,
#     `Odds Ratio` = OddsRatio,
#     `P-value` = p.value
#   ) %>%
#   gt(groupname_col = "Model") %>%
#   fmt_number(columns = c(Coefficient, `Std. Error`, `Odds Ratio`, `P-value`), decimals = 3) %>%
#   cols_align("center")



gtsave(
  data = supp_table_save,
  filename = here("outputs", "tw_Supplemental_Regression_Table.rtf")
  )

gtsave(main_table_save, here("outputs", "tw_regression_table_main.png"), vwidth = 2000, vheight = 900, expand = 5)
gtsave(supp_table_save, here("outputs", "tw_regression_table_supplement.png"), vwidth = 2000, vheight = 1600, expand = 5)

gtsave(main_table_save,here("outputs", "tw_regression_table_main.pdf"))
gtsave(supp_table_save,here("outputs", "tw_regression_table_supplement.pdf"))
```


```{r}
model_all_trimesters_together <- polr(sum_nds_category ~ 
        avg_temp_pre_conception +
        avg_temp_first_trimester +
        avg_temp_second_trimester +
        avg_temp_third_trimester +
        age + bmi + wealthindex + pregchn + season_delivery + married_bin + medlev_bin,
     data = stress_data_with_temp,
     method = "logistic",
     na.action = na.omit)


# use same tidy_polr helper function created abovve

temp_terms <- c(
  "avg_temp_pre_conception",
  "avg_temp_first_trimester",
  "avg_temp_second_trimester",
  "avg_temp_third_trimester"
)

model_all_trimesters_together_tbl <- tidy_polr(model_all_trimesters_together) %>%
  mutate(
    OR = exp(estimate),
    lo = exp(estimate - 1.96 * std.error),
    hi = exp(estimate + 1.96 * std.error),
    `OR (95% CI)` = sprintf("%.2f (%.2f, %.2f)", OR, lo, hi),
    N = nrow(model.frame(model_all_trimesters_together))
  ) %>%
  select(term, OR, lo, hi, `OR (95% CI)`, p.value, N)


model_all_trimesters_together_tbl_save <- model_all_trimesters_together_tbl %>%
  select(-c(OR, lo, hi, N)) %>%
  rename(
    Predictor = term,
    `P-value` = p.value
  ) %>%
  gt() %>%
  fmt_number(
    columns = `P-value`,
    decimals = 3
  ) %>%
  cols_align("center")



gtsave(
  data = model_all_trimesters_together_tbl_save,
  filename = here("outputs", "tw_model_all_trimesters_together_tbl_save.rtf")
  )
```

The regression analysis explores the association between wet bulb globe temperature (tw) and stress using the Sum NDS Category during different periods of pregnancy. The results indicate that:

-   **Entire Pregnancy**: Higher average daily maximum hi during the entire pregnancy is significantly associated with belonging to a higher stress group, with a one degree increase in hi associated with a 55.7% increase in the odds of belonging to a higher stress group (p = 0.004).

-   **Pre-Conception**: Higher average daily maximum hi during the pre-conception period is significantly associated with belonging to a higher stress group, with a one degree increase in hi associated with a 19.3% increase in the odds of belonging to a higher stress group (p = 0.038).

-   **Trimester-Specific Averages**:  Higher average daily maximum hi during the first trimester is significantly associated with belonging to a higher stress group, with a one degree increase in hi associated with a 35.1% increase in the odds of belonging to a higher stress group (p = 0.002). While the second trimester showed a positive association between temperature and Sum NDS Category, the effect was not significant. The third trimester average showed no effect, indicating that the temperatures nearest to the survey date appear less strongly associated with the Sum NDS Category.  





# Distributions of Data

## Distributions of negative events and NDS across domains

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Select only the `events` columns
events_columns <- crisis_dlnm_eligible %>%
  select(ends_with("events")) %>%
  select(-total_events)

events_summary <- colSums(events_columns)


# Select only the `neg` columns
neg_columns <- crisis_dlnm_eligible %>%
  select(ends_with("neg"))

# Count occurrences of numeric values in each column
neg_summary <- sapply(neg_columns, function(col) sum(as.numeric(col), na.rm = TRUE)) # warning is because function can't convert "No Event" to numeric. It makes then NAs, which for this purpose works great

# Convert to a data frame for plotting
neg_summary_df <- data.frame(
  Domain = names(neg_summary),
  Count = neg_summary
)

neg_event_labels <- c(
  crifinneg = "Financial", 
  crilegalneg = "Legal", 
  cricareerneg = "Career", 
  crirelneg = "Relationship", 
  crihomesafeneg = "Home safety", 
  crineighsafeneg = "Neighborhood safety", 
  crimedselfneg = "Self medical", 
  crimedothneg = "Other medical", 
  crihomeneg = "Home", 
  criprejneg = "Prejudice", 
  criauthneg = "Authority", 
  criothneg = "Other"
)

# Plot with the updated labels
neg_events_domain_dist <- ggplot(neg_summary_df, aes(x = reorder(Domain, -Count), y = Count)) +
  geom_bar(stat = "identity") +
  geom_col(color = "black", fill = "grey40", alpha = 1) +
  scale_x_discrete(labels = neg_event_labels) +
  labs(
    title = "(A) Total number of negatively perceived \nevents reported across domains.",
    x = "Domain",
    y = "Number of Events"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))




# Select only the `nds` columns
nds_columns <- crisis_dlnm_eligible %>%
  select(ends_with("nds")) %>%
  select(-sum_nds)

# Count occurrences of numeric values in each column
nds_summary <- sapply(nds_columns, function(col) sum(as.numeric(col), na.rm = TRUE)) # warning is because function can't convert "No Event" to numeric. It makes then NAs, which for this purpose works great

# Convert to a data frame for plotting
nds_summary_df <- data.frame(
  Domain = names(nds_summary),
  Count = nds_summary
)

nds_labels <- c(
  fin_nds = "Financial", 
  leg_nds = "Legal", 
  car_nds = "Career", 
  rel_nds = "Relationship", 
  homesf_nds = "Home safety", 
  neighsf_nds = "Neighborhood safety", 
  medself_nds = "Self medical", 
  medoth_nds = "Other medical", 
  home_nds = "Home", 
  prej_nds = "Prejudice", 
  auth_nds = "Authority", 
  oth_nds = "Other"
)

sum_nds_domain_dist <- ggplot(nds_summary_df, aes(x = reorder(Domain, -Count), y = Count)) +
  geom_bar(stat = "identity") +
  geom_col(color = "black", fill = "grey40", alpha = 1) +
  scale_x_discrete(labels = nds_labels) +
  labs(
    title = "(B) Number of participants who experienced \nat least one negative event in each domain.",
    x = "Domain",
    y = "Number of Participants"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


combined_fig <- neg_events_domain_dist + sum_nds_domain_dist +
  plot_annotation(
    title = "Figure S2. Distribution of negatively perceived events by domain",
    theme = theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5)
    )
  )








# helper to make a consistent bar panel
make_domain_bar <- function(df, labels, ylab, subtitle) {
  ggplot(df, aes(x = reorder(Domain, -Count), y = Count)) +
    geom_col(fill = "grey40", color = "black", width = 0.85) +
    scale_x_discrete(labels = labels) +
    labs(x = "Domain", y = ylab, subtitle = subtitle) +
    coord_cartesian(ylim = c(0, max(df$Count, na.rm = TRUE) * 1.08)) +
    theme_minimal(base_size = 12) +
    theme(
      plot.subtitle = element_text(face = "bold", hjust = 0, margin = margin(b = 5)),
      axis.text.x   = element_text(angle = 45, hjust = 1),
      panel.grid.major.x = element_blank(),
      panel.grid.minor   = element_blank()
    )
}

# A: total # of events per domain
pA <- make_domain_bar(
  df = neg_summary_df,
  labels = neg_event_labels,
  ylab = "Number of Events",
  subtitle = "Total number of negatively perceived \nevents reported across domains"
)

# B: # of participants with ≥1 event per domain
pB <- make_domain_bar(
  df = nds_summary_df,
  labels = nds_labels,
  ylab = "Number of Participants",
  subtitle = "Participants who experienced at least one \nnegative event in each domain"
)

# combine with A/B tags; no overall title (use manuscript caption instead)
combined_fig <- (pA + pB) +
  plot_annotation(
    tag_levels = "A",
    theme = theme(
      plot.tag = element_text(face = "bold", size = 12),
      plot.tag.position = c(0.01, 0.98)  # top-left
    )
  )

combined_fig


ggsave(here("outputs", "figureS2_neg_event_domains.png"),
       combined_fig, 
       width = 9, 
       height = 5, 
       dpi = 400)



```

## Distribution of Sum NDS across participants 

```{r, echo=FALSE, warning=FALSE, message=FALSE}


# Bar for the sum NDS score
nds_count_data <- crisis_dlnm_eligible %>%
  count(sum_nds)


sample_stats <- crisis_dlnm_eligible %>%
  summarise(
    n = n(),
    median = median(sum_nds, na.rm = TRUE),
    q1 = quantile(sum_nds, 0.25, na.rm = TRUE),
    q3 = quantile(sum_nds, 0.75, na.rm = TRUE)
  ) %>%
  mutate(
    stats_label = paste0("N = ", n, "\nMedian = ", round(median, 2), "\nIQR = ", round(q1, 1), "–", round(q3, 1))
  )


## 4) Headroom for labels + a safe place for the stats box
y_max      <- max(nds_count_data$n, na.rm = TRUE)
x_min      <- min(nds_count_data$sum_nds, na.rm = TRUE)
x_max      <- max(nds_count_data$sum_nds, na.rm = TRUE)
x_statspos <- x_min + 0.75*(x_max - x_min)  # ~right side
y_statspos <- y_max * 0.70

## 5) Plot (no baked title; panel title will live in the manuscript caption)
nds_dist_plot <- ggplot(nds_count_data, aes(sum_nds, n)) +
  geom_col(fill = "grey40", color = "black", width = 0.85) +
  geom_text(aes(label = n), vjust = -0.4, size = 3) +
  annotate("label",
           x = x_statspos, y = y_statspos,
           label = sample_stats$stats_label,
           label.size = 0, fill = "white", alpha = 0.9) +
  scale_x_continuous(breaks = seq(x_min, x_max, by = 1)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +  # top headroom
  labs(x = "Sum NDS", y = "Participants") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(10, 15, 10, 10)
  )

nds_dist_plot




ggsave(
  here("outputs", "figureS1_nds_dist.png"),
  nds_dist_plot,
  width = 8,
  height = 6,
  dpi = 300
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
crisis_recode %>%
  count(sum_nds_category) %>%
  ggplot(aes(x = sum_nds_category, y = n))  +
  geom_col(fill = "steelblue", color = "black") +
  geom_text(aes(label = n), vjust = -0.5) +
  theme_minimal() +
  labs(title = "Distribution of Sum NDS Category", x = "Sum NDS Category", y = "Participants")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
print("Summary of Sum NDS data")
summary(crisis_dlnm_eligible$sum_nds)
```


# Participant and exposure characteristics table 

```{r}
# Packages
library(tidyverse)
library(gtsummary)
library(labelled)
library(forcats)
library(janitor)
library(knitr)
library(kableExtra)
library(gt)

# Temperature: compute rowwise Min/Max tw across all lag columns

# Number of pre- and post-conception days
max_gestation_days <- 210
pre_conception_days <- 30

# ---- helper that also tracks completeness and drops sparse rows ----
compute_participant_mean <- function(era5_data, metric, max_gestation_days = 210, pre_conception_days = 30,
                                     min_prop_days = 0.8) { # min prop days is not necessary for our data, which is 100% complete 
  temp_df <- era5_data %>%
    select(community, date, !!sym(metric)) %>%
    rename(temp_metric = !!sym(metric)) %>%
    distinct()

  total_days <- pre_conception_days + max_gestation_days + 1  # inclusive of day_0

  lags_long <- crisis_dlnm_eligible %>%
    select(mstudyid, vname, date_conception) %>%
    rowwise() %>%
    mutate(day_relative = list(-pre_conception_days:max_gestation_days),
           date = list(date_conception + day_relative)) %>%
    unnest(c(day_relative, date)) %>%
    left_join(temp_df, by = c("vname" = "community", "date" = "date"))

  lags_long %>%
    group_by(mstudyid) %>%
    summarise(
      n_days = sum(!is.na(temp_metric)),
      prop_days = n_days / total_days,
      mean_temp = if_else(n_days > 0, mean(temp_metric, na.rm = TRUE), NA_real_),
      .groups = "drop"
    ) %>%
    filter(prop_days >= min_prop_days) %>%                         # keep well-observed pregnancies
    rename(!!paste0("mean_", metric) := mean_temp)
}

# run for both metrics (80% completeness threshold; change if you prefer)
mean_max <- compute_participant_mean(era5, "max_tw", min_prop_days = 0.8)
mean_min <- compute_participant_mean(era5, "min_tw", min_prop_days = 0.8)

# join + factor prep
crisis_part <- crisis_dlnm_eligible %>%
  left_join(mean_max, by = "mstudyid") %>%
  left_join(mean_min, by = "mstudyid") %>%
  mutate(
    ses_num   = as.numeric(wealthindex),
    married_f = labelled::to_factor(married) %>% forcats::fct_drop(),
    medlev_f  = labelled::to_factor(medlev)  %>% forcats::fct_drop(),
    crisys_cat_f = factor(sum_nds_category, ordered = TRUE) %>% forcats::fct_drop(),
    season_delivery = as.factor(season_delivery)
  )

# build table
tbl1 <- crisis_part %>%
  select(
    age, bmi, pregchn,season_delivery, ses_num,
    married_f, medlev_f,
    sum_nds, crisys_cat_f,
    mean_min_tw, mean_max_tw
  ) %>%
  gtsummary::tbl_summary(
    type = list(all_continuous() ~ "continuous", all_categorical() ~ "categorical"),
    statistic = list(all_continuous() ~ "{median} ({p25}, {p75})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_continuous() ~ 1, all_categorical() ~ 0),
    label = list(
      age ~ "Age (years)",
      bmi ~ "BMI (kg/m²)",
      pregchn ~ "Number of Previous Pregnancies",
      season_delivery ~ "Season of Delivery",
      ses_num ~ "SES (1–5, 1 = least poor and 5 = very poor)",
      married_f ~ "Marital status",
      medlev_f ~ "Education level",
      sum_nds ~ "Sum NDS score",
      crisys_cat_f ~ "Sum NDS category",
      mean_min_tw ~ "Mean minimum tw (°C)",
      mean_max_tw ~ "Mean maximum tw (°C)"
    ),
    missing = "ifany"
  ) %>%
  gtsummary::bold_labels() %>%
  gtsummary::modify_header(
    label ~ "**Characteristic**",
    all_stat_cols() ~ "**Median (IQR) or n (%)**"
  ) %>%
  gtsummary::as_gt() %>%
  gt::tab_options(
    table.font.size = gt::px(12),
    table.width = gt::pct(60),
    heading.align = "left",
    data_row.padding = gt::px(4),
    column_labels.font.weight = "bold"
  )



gtsave(
  data = tbl1,
  filename = here("outputs", "table1_participant_exposure.rtf")
)



gtsave(tbl1, filename = here("outputs", "table1_participant_exposure.png"))  # for figures
gtsave(tbl1, filename =here("outputs", "table1_participant_exposure.pdf"))             # for manuscript

```



# Compare characteristics of dlnm study participants to broader graphs cohort

```{r, echo=FALSE, message=FALSE, warning=FALSE}

min(crisis_dlnm_eligible$survey_date)
max(crisis_dlnm_eligible$survey_date)

# calculate midpoint of stress data so can calculate age comparison between stress sample and wider GRAPHS cohort
start_date <- as.Date("2014-09-09")
end_date   <- as.Date("2015-11-02")

midpoint_date <- start_date + as.numeric(difftime(end_date, start_date, units = "days")) / 2
midpoint_date


graphs_covariates <- read_csv(here("data", "alldata_clean_for_bw_19May08.csv")) %>%
  select(mstudyid, 
         pregchn, # total number of lifetime pregancies 
         dob,
         medlev,  # educational level of mother
         married,
         ses,
         asset_index,
         is_poor,
         bmi)


not_included_sample <- graphs_covariates %>% 
  filter(!mstudyid %in% c(crisis_dlnm_eligible$mstudyid)) %>%
  mutate(
    dob = mdy(dob),
    age_at_reference = time_length(interval(dob, midpoint_date), "years")
  ) %>%
  mutate(group = "Not included in analysis (wider GRAPHS cohort)") %>%
  mutate(
    married = ifelse(is.na(married), 7, married),
    medlev = ifelse(is.na(medlev), 7, medlev),
    ses = ifelse(is.na(ses), "Unknown", ses),
    ses = as.factor(ses),
    
    married_f = factor(married,
                       levels = c(1, 2, 3, 4, 5, 6, 7),
                       labels = c("Married", "Living together", "Widowed", "Divorced", "Separated", "Single", "Unknown")),

    medlev_f = factor(
      medlev,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c(
        "None", "Primary", "JHS", "SHS",
        "Post-middle training", "Post-secondary training", "Unknown"
      )
    ),
    
  ) 

library(gtsummary)

table_graphs <- not_included_sample %>%
  select(
    age_at_reference,
    bmi,
    pregchn,
    ses,
    married_f,
    medlev_f
  ) %>%
  tbl_summary(
    type = list(
      all_continuous() ~ "continuous",
      all_categorical() ~ "categorical"
    ),
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_continuous() ~ 1,
      all_categorical() ~ 0
    ),
    label = list(
      age_at_reference ~ "Age (years)",
      bmi ~ "BMI (kg/m²)",
      pregchn ~ "Number of Previous Pregnancies",
      ses ~ "SES (1–5)",
      married_f ~ "Marital Status",
      medlev_f ~ "Education Level"
    ),
    missing = "ifany"
  ) %>%
  bold_labels() %>%
  modify_header(
    label ~ "**Characteristic**",
    all_stat_cols() ~ "**Mean (SD) or n (%)**"
  )

table_graphs_gt <- table_graphs %>% as_gt()

gtsave(
  data = table_graphs_gt,
  filename = here("outputs", "tables1_graphs.rtf")
)



included <- crisis_dlnm_eligible %>%
  mutate(
    group = "Included in analysis",
    wealthindex = as.factor(wealthindex),

    married_f = factor(
      married,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c("Married", "Living together", "Widowed",
                 "Divorced", "Separated", "Single", "Unknown")
    ),

    medlev_f = factor(
      medlev,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c("None", "Primary", "JHS", "SHS",
                 "Post-middle training", "Post-secondary training", "Unknown")
    )
  ) %>%
  select(
    group,
    age,
    bmi,
    pregchn,
    wealthindex,
    married_f,
    medlev_f
  )


excluded <- not_included_sample %>%
  rename(age = age_at_reference,
         wealthindex = ses) %>%
  mutate(group = "zExcluded from analysis (wider GRAPHS cohort)") %>%
  select(
    group,
    age,
    bmi,
    pregchn,
    wealthindex,
    married_f,
    medlev_f
  )


comparison_df <- bind_rows(included, excluded)

library(gtsummary)

combined_table <- comparison_df %>%
  tbl_summary(
    by = group,
    type = list(
      all_continuous() ~ "continuous",
      all_categorical() ~ "categorical"
    ),
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "ifany",
    label = list(
      age ~ "Age (years)",
      bmi ~ "BMI (kg/m²)",
      pregchn ~ "Number of Previous Pregnancies",
      wealthindex ~ "SES (1–5)",
      married_f ~ "Marital Status",
      medlev_f ~ "Education Level"
    )
  ) %>%
  bold_labels() %>%
  modify_header(
    label ~ "**Characteristic**",
    all_stat_cols() ~ "**Mean (SD) or n (%)**"
  )

combined_table_gt <- combined_table %>% as_gt()



gtsave(
  data = combined_table_gt,
  filename = here("outputs", "table_s1_graphs_comparison.rtf")
)


# summary_table <- combined_samples %>%
#   group_by(group) %>%
#   summarise(
#     n = n(),
#     `Age (years)` = sprintf("%.1f (%.1f)", mean(age_at_reference, na.rm = TRUE), sd(age_at_reference, na.rm = TRUE)),
#     `BMI (kg/m²)` = sprintf("%.1f (%.1f)", mean(bmi, na.rm = TRUE), sd(bmi, na.rm = TRUE)),
#     `SES (1–5)` = sprintf("%.2f (%.2f)", mean(ses, na.rm = TRUE), sd(ses, na.rm = TRUE)),
#     `Pregnancies (n)` = sprintf("%.1f (%.1f)", mean(pregchn, na.rm = TRUE), sd(pregchn, na.rm = TRUE)),
#     `Married (%)` = sprintf("%.0f%%", 100 * mean(married == 1, na.rm = TRUE))
#   )
# 
# table_s1 <- summary_table %>%
#   t() %>%
#   as.data.frame() %>%
#   rownames_to_column("Characteristic")
# 
# names(table_s1) <- c("Characteristic", "Included in analysis", "Not included in analysis (wider GRAPHS cohort)")
# 
# table_s1 <- table_s1[-1, ]  # remove the 'group' row
# 
# 
# table_graphs_comp_s1 <- table_s1 %>%
#   kable(
#     format = "html",
#     align = "lcc",
#     row.names = FALSE   # 👈 This removes the numbered row names
#   ) %>%
#   kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE) %>%
#   kable_styling(font_size = 15) %>%
#   footnote(
#     general_title = "",
#     footnote_as_chunk = TRUE,
#     escape = FALSE
#   )
# 
# 
# 
# # Save as high-resolution PNG for Word
# save_kable(
#   table_graphs_comp_s1,
#   file = here("outputs", "tables1_graphs_crysis_comp.png"),
#   zoom = 3,             # higher = higher DPI
#   vwidth = 800,        # width in pixels (default ~992)
#   vheight = 500,        # height in pixels
#   density = 300         # DPI for high-res output
# )
# 
# # Save as PDF
# save_kable(table_graphs_comp_s1, file = here("outputs", "tables1_graphs_crysis_comp.pdf"), self_contained = TRUE)
```





